Importance sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-04-15 draft_p2.tex
Fix typos

% Motivation
In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time unit $tau_c$.
However, In many applications, this duration is much smaller than the timescale of interest.
Consider for insatnce the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.

@@

%% Objectif des algo.
In this section we discuss the application of \textbf{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example
of the flow presented in section~\ref{sec:test_flow}.
The purpose of rare events algorithms is to make it possible to sample rare events for a computational cost well inferior to their return time.
In sections~\ref{sec:ams} and~\ref{sec:gktl} we consider the \acl{ams} and \acl{gktl} algorithms, respectively.
Both algorithms rely on the simulation of an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{1\leq n \leq N, 0\leq t T_a}$, where $\{\mathbf{x}(t)\}_{0\leq t T_a}$ is a formal notation for a trajectory
of duration $T_a$ in the phase space describing the dynamical system.
In this case the dynamical system is the flow described in section~\ref{sec:test_flow}.
Over the iterations of the algorithm, trajectories are duplicated or discarded from the ensemble according to precise selection rules that bias the sampling towards the rare events of interest.
Such selection rules are designed in such a way that the sampling bias is known at each iteration of the algorithm.
In other words, generated trajectories are given a statistical weight, which knowledge allows for the computation of both their probability and expectation values of any trajectory observables.

@@

% Application to turbulent flow past an obstacle
Can the \ac{tams} achieve similar results when applied to complex chaotic dynamics, such as turbulent flows around objects ?
To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
Our aim is to sample trajectories that display extreme fluctuations of the drag $f_d$ acting on the square obstacle.
However, by contrast with a simple \ac{ou} process, the phase space is higly-dimensional and the dynamics complex.
As a consequence the choice of the score function $xi (\mathbf{x}(t),t)$ is not so clear.
In the following we use the \ac{tams} with $\xi \equiv f_d$, which is the most simple choice for the score function.
For such complex dynamics however, there is no reasons for the drag itself to be a good measure of how far is a given state to regions of phase space where the drag is extreme.
Unfortunately, the choice of better score functions rely on a deep understanding of the physics of the rare events of interest, which is precisely what we are trying to achieve.

>>>>

% Motivation
In section~\ref{sec:instantaneous_drag}, we discussed the phenomenology of rare events corresponding to extremely high values of the drag acting on the square obstacle mounted in the flow described in section~\ref{sec:test_flow}.
In particular, it was pointed out that such extreme drag fluctuations have a lifetime of, roughly, one correlation time unit $\tau_c$.
However, In many applications, this duration is much smaller than the timescale of interest.
Consider for instance the interaction of a deformable structure with a turbulent flow: the typical response time may be much larger than the lifetime of drag fluctuations.

@@

%% Objectif des algo.
In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example
of the flow presented in section~\ref{sec:test_flow}.
The purpose of rare events algorithms is to make it possible to sample rare events for a computational cost well inferior to their return time.
In sections~\ref{sec:ams} and~\ref{sec:gktl} we consider the \acl{ams} and \acl{gktl} algorithms, respectively.
Both algorithms rely on the simulation of an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{1\leq n \leq N, 0\leq t T_a}$, where $\{\mathbf{x}(t)\}_{0\leq t T_a}$ is a formal notation for a trajectory
of duration $T_a$ in the phase space describing the dynamical system.
In this case the dynamical system is the flow described in section~\ref{sec:test_flow}.
Over the iterations of the algorithm, trajectories are duplicated or discarded from the ensemble according to precise selection rules that bias the sampling towards the rare events of interest.
Such selection rules are designed in such a way that the sampling bias is known at each iteration of the algorithm.
In other words, generated trajectories are given a statistical weight, which knowledge allows for the computation of both their probability and expectation values of any trajectory observables.

@@

% Application to turbulent flow past an obstacle
Can the \ac{tams} achieve similar results when applied to complex chaotic dynamics, such as turbulent flows around objects ?
To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
Our aim is to sample trajectories that display extreme fluctuations of the drag $f_d$ acting on the square obstacle.
However, by contrast with a simple \ac{ou} process, the phase space is higly-dimensional and the dynamics complex.
As a consequence the choice of the score function $\xi (\mathbf{x}(t),t)$ is not so clear.
In the following we use the \ac{tams} with $\xi \equiv f_d$, which is the most simple choice for the score function.
For such complex dynamics however, there is no reasons for the drag itself to be a good measure of how far is a given state to regions of phase space where the drag is extreme.
Unfortunately, the choice of better score functions rely on a deep understanding of the physics of the rare events of interest, which is precisely what we are trying to achieve.
